---
title: "Using stan without heirarchical model"
author: "Joe Skufca"
date:  "2021-12-23"
output: 
    html_notebook:
       toc: true
---

This experiment intends to analyze data from all farm as a single group classes, not incorporating the "herd" component. It will predict class for each animal.  It will allow us to explore (a bit, with the affect of "prior").   The dataset of focus here is the "goat" dataset.

### Setup

```{r setup}
options(digits = 3)
knitr::opts_chunk$set( fig.width = 6,highlight=T)

library(tidyverse)
library(janitor)
library(PriorGen)


library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores(logical = FALSE))

library(bayesplot)
library(here)
```

## Example 1

We will analyze the goat data, but without a heirarchical model (so no need to include which farm).

### Data

Load data.  We eliminate the one sample that was questionable.


```{r}
goat_file=here("data","goat_results.xlsx")

goat = readxl::read_excel(goat_file, sheet = "Sheet1") %>%
     clean_names() %>%
     select(sample_number, elisa, pcr, culture) %>%
     filter(elisa  %in% c("p", "n")) %>%
     filter(culture %in% c("p", "n")) %>%
     mutate(across(elisa:culture, ~ (.x == "p") + 1)) 


```

### Assemble Stan model data

In addition to the data, we need information for prior assumptions regarding sensitivity and specificity.

If we start with a non-informative prior:



```{r}
data2= list(I=316,
            J=3,
            y=as.matrix(goat %>% select(2:4)),
            alpha_sens = c(1,1,1),
            beta_sens = c(1,1,1),
            alpha_spec = c(1,1,1),
            beta_spec= c(1,1,1)
                    )

#            alpha_sens = c(167,167,167),
#            beta_sens = c(43,43,43),
#            alpha_spec = c(43,43,43),
#            beta_spec= c(5.62,5.62,5.62)
```


### fitting the model

```{r}
fit2=stan(file=here("src","multi_test_spec_sens_prev.stan"),data=data2)
```

###  Visualizing results

```{r}
posterior=as.matrix(fit2)
```


Prevalence

```{r}
plot_title <- ggtitle("Posterior distributions",
                      "with medians and 80% intervals")
mcmc_areas(posterior,prob=.8,pars="pi[2]")


```

Specificity:
```{r}
p=mcmc_areas(posterior,prob=.8,regex_pars=c("spec"))
p+scale_y_discrete(labels=c("spec_ELISA","spec_pcr","spec_culture"))
```


Sensitivity:
```{r}
p=mcmc_areas(posterior,prob=.8,regex_pars=c("sens"))
p+scale_y_discrete(labels=c("sens_ELISA","sens_pcr","sens_culture"))
```
### helper function

```{r}
fitandplot=function (newdata) {
     fit2=stan(file=here("src","multi_test_spec_sens_prev.stan"),data=newdata)
     posterior=as.matrix(fit2)
     plot_title <- ggtitle("Posterior distributions",
                      "with medians and 80% intervals")
p=mcmc_areas(posterior,prob=.8,pars="pi[2]")
print(p)
p=mcmc_areas(posterior,prob=.8,regex_pars=c("spec"))
p=p+scale_y_discrete(labels=c("spec_ELISA","spec_pcr","spec_culture"))
print(p)
p=mcmc_areas(posterior,prob=.8,regex_pars=c("sens"))
p=p+scale_y_discrete(labels=c("sens_ELISA","sens_pcr","sens_culture"))
print(p)
fit2
}
```


### Other Priors

Taking values from Ontario paper:

```{r}
data3= list(I=316,
            J=3,
            y=as.matrix(goat %>% select(2:4)),
            alpha_sens = c( 4.33,   2.41,    4.98),
            beta_sens =  c( 8.77,   4.29,    6.96),
            alpha_spec = c(99.7 , 107.2 , 2291.17),
            beta_spec=   c( 6.19,   3.17,    6.74)
                    )

```

```{r}
fit3=fitandplot(data3)
```

#### helper to compare priors

```{r}
#define range
plot_priors=function(a,b) {
x = seq(0,1, length=1000)

p <- ggfortify::ggdistribution(dbeta, x, shape1 = a[1],shape2=b[1], colour = 'blue')
p <- ggfortify::ggdistribution(dbeta, x, shape1 = a[2],shape2=b[2], colour = 'green', p = p)
p= ggfortify::ggdistribution(dbeta, x, shape1 = a[3],shape2=b[3], colour = 'red', p = p)
p=p+ggtitle("ELISA - blue, PCR - green, Culture - red")
print(p)
}
```

```{r}
plot_priors( c(99.7 , 107.2 , 2291.17),
             c( 6.19,   3.17,    6.74))
```

```{r}
mcmc_intervals(as.matrix(fit3),regex_pars="^z\\[[1-3][0-9],2\\]")
```

### Let's suppose we have really strong belief in the ELISA (sensitivity)

```{r}
data4= list(I=316,
            J=3,
            y=as.matrix(goat %>% select(2:4)),
            alpha_sens = c( 2224.33,   2.41,    4.98),
            beta_sens =  c( 8.77,   4.29,    6.96),
            alpha_spec = c(99.7 , 107.2 , 2291.17),
            beta_spec=   c( 6.19,   3.17,    6.74)
                    )

```

```{r}
fit4=fitandplot(data4)
```


