---
title: "Exp 14:Using stan with heirarchical model"
author: "Joe Skufca"
date: "2024-02-09"
output:
  html_document:
    toc: yes
    df_print: paged
  powerpoint_presentation:
    slide_level: 3
  word_document:
    toc: yes
---

Experiment 14 is a modification to the technique applied in experiment 13, but uses a different approach.

We start with the collected data from the grant funded study and
uses that data as the basis for a model where we will add a "prospective" farm.

The intention is to be able to make predictions for HIP and CWHP for a farm 
that lyes outside the original protocol tested.   Typically, we image this will be
"new" data, but likely only from one test.

* We use a two-latent class model, which assumes each animal is classes as either "infected" or "non-infected."
* We make estimates of within herd prevalence for each herd.
* For the "new" herd, we will append whatever test results are available.
* We will add (new) indicator variables that allow for differentiating between "valid" (actual) and "invalid" (not actually taken) tests.


> This approach for missing data is connected with `heir_mult_test_B_OOS_2.stan`



* Prior for herd level prevelence was initially taken as $Beta(1.55,2.66)$, but is modified below.  



## Setup

We load anticipated packages and set rstan options.

```{r setup}
options(digits = 3)
knitr::opts_chunk$set( highlight=T,message=FALSE, warning=FALSE,echo=FALSE)

library(tidyverse)
library(patchwork)
library(janitor)
library(PriorGen)
library(ggpmisc)

library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores(logical = FALSE))

library(bayesplot)
library(here)
library(tidybayes)
library(shinystan)
library(DT)
library(flextable)
```

## Preparing the goat data

We will analyze the goat data, but using a hierarchical model (with each farm treated as a herd).

### Historical (baseline) Data

Load data.  We eliminate the one sample that was questionable.


```{r message=FALSE, warning=FALSE, results="asis"}
goat_file=here("data","goat_results.xlsx")

#goat = readxl::read_excel(goat_file, sheet = "Sheet1") %>%
goat = readxl::read_excel(goat_file, sheet = "surrogate_data_1") %>%
     clean_names()%>% mutate(sex=str_to_lower(sex)) %>%
     mutate(sex=case_when(
          str_starts(sex,"f") ~ "f",
          TRUE ~ "m"
     )) %>%
     select(sample_number,sex,farm_number,  elisa, pcr, culture) %>%
     filter(elisa  %in% c("p", "n")) %>%
     filter(culture %in% c("p", "n")) %>%
     mutate(across(elisa:culture, ~ (.x == "p") + 0)) %>%
     mutate(valid_elisa=1,valid_pcr=1,valid_culture=1)

#gtsummary::tbl_summary(goat,include=c(sex,elisa,pcr,culture))

```

### Additional data from new farm

Similar to the goat set above, we will add hypothetical data, using -1 for missing results.

Hypothetical - test of 20 goats with ELISA, but no other tests performed.


```{r}
this_elisa=c(rep(0,20))  #  20 total, all negative
#this_elisa=c(rep(0,19),1)  #  20 total, 1 positive
#this_elisa=c(0, 0, 0, 0, 0, 0)  #  6, all negative
#this_elisa=c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)  #  11, all negative
#this_elisa=rep(0,30)  #  30, all negative
s=length(this_elisa)




s_max=max(goat$sample_number)
f_max=max(goat$farm_number)

goat_new=tibble(sample_number=1:s + s_max,
                farm_number=f_max+1,
                elisa=this_elisa,  # simulated real samples
                pcr=0,  # filler 
                culture=0,
                valid_elisa=1,
                valid_pcr=0,
                valid_culture=0)        # filler to square up the data frame

goat_s=goat %>% bind_rows(goat_new)
```





### Assemble Stan model data

I include here a breakdown table by farm, as it is part of assembling the data structure required to input to the stan code. This table will help us to understand why certain farms are deemed to be higher likelyhood of being infected as well as differing levels of within farm prevelance.

```{r}
goat_sum=goat_s %>% rowid_to_column("id") %>% group_by(farm_number) %>%
     summarise(i_s=first(id),i_e=last(id),n=n(),across(elisa:culture,sum))
goat_sum %>% flextable() 
```

The provided information provides the measured data and its structure, along with several parameters and hyperparameters that define the prior distributions.



### Prior information

We use the `PriorGen` package to assist with developing hyperparameters.  The wrapper functions created here allow the text output from that packaage to be used programmatically.

```{r }
my_findbeta=function(...){
     pattern <- "[0-9.]+"
     a=capture.output(findbeta(...))
     b=str_match_all(a[1],pattern) %>% unlist() %>% as.numeric()
     b=b[2:3]
     b
}

my_findbetamupsi=function(...){
     pattern <- "[0-9.]+"
     a=capture.output(findbetamupsi(...))
     b=str_match_all(a[1],pattern) %>% unlist() %>% as.numeric()
     b=b[2:3]
     g=str_match_all(a[2],pattern) %>% unlist() %>% as.numeric()
     g=g[2:3]
     c(b,g) #output
}

```

#### Hyperparameter for priors

For several of the hyperparameters, we will use Beta or Gamma distributions as priors, along with expert knowledge to develop informative priors.  We provide several examples, as templates, (with blank informationation), then continue with simply showing the completed information.

##### Herd True Prevelance

>For HTP hyperparameters: Based on expert opinion, we suspect that the most
>likely value for the fraction of farms with Johnes present in the goat flock is
>about `v1`, and we are **95%** certain that the fraction of infected farms is
>less that `v2`.

```{r}
htp1=.3
htp2=.67
htp_hyp=my_findbeta(themode = htp1,percentile.value = htp2,lower.v = TRUE)
ggfortify::ggdistribution(dbeta,seq(0,1,.005),shape1=htp_hyp[1],shape2=htp_hyp[2])+theme_minimal()
```

For HTP hyperparameters: Based on expert opinion, we suspect that the most likely value for the fraction of farms with Johnes present in the goat flock is about `r formattable::percent(htp1)`, and we are **95%** certain that the fraction of infected farms is less that `r formattable::percent(htp2)`.


### Within Herd Prevalence (WHP)

Within herd prevalence prior information is captured by both a mean value and a description of the variance, with hyperparameters $\mu$ and $\psi$ drawn from a Beta and Gamma distribution respectivelvy. 
To generate the hyperparameters for $\mu$ and $\psi$ so that we can prescribe priors for within herd prevalence, we complete the  portions of this sentence with expert opinion based on area knoweledge:

```{r}
v1=.2 
v2=.95
v3= .33
v4=.9
v5=.5
v6=.7

mupsi_hyp=my_findbetamupsi(themean=v1, percentile=v2, percentile.value=v3,
 lower.v=TRUE,
psi.percentile=v4, percentile.median=v5, percentile95value=v6)

```

> Consider herds in this region of study, and consider those farms for which it is known that
> Johnes is present in that herd.  The fraction of animals on that farm that will be infected
> would be: the **average** value is thought to be `r formattable::percent(v1)`  and we are 
> `r formattable::percent(v2)` confident that **average value** it is not more than `r formattable::percent(v3)`.
> Rather than the *average value*, we may estimate an upper bound, such that 
> `r formattable::percent(v4)` of the infected farm would probably have an infection rate below
> `r formattable::percent(v5)`, and we are 99% certain that they would not exceed `r formattable::percent(v6)`.

### Sensitivity and Specificity

We also need prior for sensitivity and specificity for each of the three tests.  One of the challenges here is that we are nominally using the test as measures of the presense of infection.  The tests used here have a good bit of independence (using different biological signals to produce a test outcome) which is an assumption of the statistical model that is used.  However, the infection progression in the body is such that not all of the signals would necessarily be present at the same time, such that the assumption that a "diseased" goat should test positive on all test is not a precise description of the biology.  As such, the effect (in an operational setting) is that the test would typically show lower accuracy than they would compared to the "gold standard" for that particular type of test.  (For example - the manufacturer statements about accuracty for PCR are in comparison to a gold standard PCR test - equivalently does the PCR test being used detect DNA material when bacterial DNA material is actually present in the blood at a level above some known minimum.  The host might be shedding bacteria in the fecal matter at a level which can be cultured even when not detectable in the blood as this shedding results from a different biological process.)

To allow for this reduced accuracy within our bayesian analysis, we widen our assumed prior distributions, making them less informative.  We use the modes as suggested, but using 95% intervals that allow for a much broader range of error, allowing that the data might have stronger influence of the posterior distribution.

For each of these parameters, we use a template that follows this exemplar pattern:

For parameter xxx - use mode of (   ), with 95% probability that is value is above (   )%

```{r}
elisa_v=c(1,.33,.985,.9)#my changes on specificity
pcr_v=c(.956,.33,1,.9) #my changes on specificity
culture_v=c(.231,.1,.99,.9) #my changes on specificity
elisa_sens=my_findbeta(themode = elisa_v[1],percentile.value = elisa_v[2],lower.v = FALSE)
elisa_spec=my_findbeta(themode = elisa_v[3],percentile.value = elisa_v[4],lower.v = FALSE)
pcr_sens=my_findbeta(themode = pcr_v[1],percentile.value = pcr_v[2],lower.v = FALSE)
pcr_spec=my_findbeta(themode = pcr_v[3],percentile.value = pcr_v[4],lower.v = FALSE)
culture_sens=my_findbeta(themode = culture_v[1],percentile.value = culture_v[2],lower.v = FALSE)
culture_spec=my_findbeta(themode = culture_v[3],percentile.value = culture_v[4],lower.v = FALSE)
```

* Elisa sens: use mode of `r formattable::percent(elisa_v[1])`, with 95% probability that is value is above `r formattable::percent(elisa_v[2])`
* Elisa spec: use mode of `r formattable::percent(elisa_v[3])`, with 95% probability that is value is above `r formattable::percent(elisa_v[4])`
* pcr sens: use mode of `r formattable::percent(pcr_v[1])`, with 95% probability that is value is above `r formattable::percent(pcr_v[2])`
* pcr spec: use mode of `r formattable::percent(pcr_v[3])`, with 95% probability that is value is above `r formattable::percent(pcr_v[4])`
* culture sens: use mode of `r formattable::percent(culture_v[1])`, with 95% probability that is value is above `r formattable::percent(culture_v[2])`
* culture spec: use mode of `r formattable::percent(culture_v[3])`, with 95% probability that is value is above `r formattable::percent(culture_v[4])`


```{r}

data5= list(F=nrow(goat_sum),
            I=nrow(goat_s),
            I_s=goat_sum$i_s,
            I_e=goat_sum$i_e,
            J=3,
            y=as.matrix(goat_s %>% select(elisa:culture)),
            v=as.matrix(goat_s %>% select(valid_elisa:valid_culture)),
            alpha_sens = c( elisa_sens[1],   pcr_sens[1],    culture_sens[1]),
            beta_sens =  c( elisa_sens[2],   pcr_sens[2],    culture_sens[2]),
            alpha_spec = c( elisa_spec[1],   pcr_spec[1],    culture_spec[1]),
            beta_spec=   c( elisa_spec[2],   pcr_spec[2],    culture_spec[2]),
            htp_alpha=htp_hyp[1],
            htp_beta=htp_hyp[2],
            mu_alpha=mupsi_hyp[1],
            mu_beta=mupsi_hyp[2],
            psi_a=mupsi_hyp[3],
            psi_b=mupsi_hyp[4]
                    )

# Parameters from Ontario 2011 paper, as reference
            # alpha_sens = c( 4.33,   2.41,    4.98),
            # beta_sens =  c( 8.77,   4.29,    6.96),
            # alpha_spec = c(99.7 , 107.2 , 2291.17),
            # beta_spec=   c( 6.19,   3.17,    6.74),

```


### fitting the model

This model takes a bit of time to run, depending upon the power of your machine.  On my office machine, with model compiled, I can compute a new simulation (new parameters) usually in about a minute.
```{r}
fit5=stan(file=here("src","heir_multi_test_B_OOS_2.stan"),data=data5,
          warmup=200,chains=12,iter=400)
```

###  Visualizing results

Primarily, here, I will be showing results with graphs, using a bit of words to describe.  

The plots will all be posterior distribution of various parameter values.

```{r}
posterior=as.matrix(fit5)
```


### Herd True Prevalence

Herd True Prevalence (HTP) is the fraction of herd with at least one infected animal.

```{r}
plot_title <- ggtitle("Herd Level Prevalence",
                      "with medians and 80% intervals")
mcmc_areas(posterior,prob=.8,pars="HTP") + plot_title
#ggsave(filename=here("figs","HTP_goat.png"), width=10,height=8,units="cm"   )
```
For a particular parameter, we can ask for detailed summary information in many different forms.  Below I provide a point estimate (median) along with an 90% hdi interval (highest density interval) estimate.   The median value indicates that about 1/3 of the farms are infected.
```{r}

fit5 %>% gather_draws(HTP) %>% median_hdi(.width=.9) %>% flextable()
```

As this mean value checks (somewhat) with the fraction of herds that had positive test results, we imagine that maybe the model is doing something that we expect.

### Sensitivity and Specificity

A critical goal to this grant, as well as to this analysis overall is ... how accurate are these tests.  Sensitivity and Specificity estimates are an *output* of the model, though we have discussed that they are affected by the assumed prior distributions.  For the analysis shown, the posterior distributions for these measures:

Specificity is very high, though ELISA is much worse than the others.

```{r }
plot_title <- ggtitle("Specificity (Goats)")
#ggsave(filename=here("figs","spec_goat.png"), width=10,height=8,units="cm"   )
mcmc_areas(posterior,regex_pars=c("spec"))+
     scale_y_discrete(labels=c("spec_elisa","spec_pcr","spec_cult"))+plot_title


#ggsave(filename=here("figs","spec_goat.png"), width=7.5,height=5,units="cm"   )
```

### Sensitivity

```{r}
plot_title <- ggtitle("Sensitivy (Goats)")
mcmc_areas(posterior,regex_pars=c("sens"))+
     scale_y_discrete(labels=c("sens_elisa","sens_pcr","sens_cult"))+plot_title

#ggsave(filename=here("figs","sens_goat.png"), width=7.5,height=5,units="cm"   )
```


```{r}
fit5 %>% gather_draws(spec[i],sens[i]) %>% median_hdi(.width=.9) %>% 
     arrange(.variable) %>%  flextable()
```

### Herd Infection Probability

The way I formulated the model allows me to compute a parameter which I call _herd infection probability_ (HIP), which give the probability that a particular herd has at least one infected animal in the group of animals tested.  (Note the care of wording.  The model is not aware of the total number of animals on the farm, which would be necessary to actually perform this calculation.)

```{r fig.height=5, include=FALSE}
mcmc_areas(posterior,pars=vars(param_range("HIP",c(1:4))))
```







```{r}
new=set_flextable_defaults(digits=3)
dfab=fit5 %>% gather_draws(HIP[i]) %>% median_hdi(.width=.9)


dfab %>% mutate(across(3:6,~ round(.x,digits=3))) %>% flextable() 
```

It is informative to merge these median values with the farm-level test results to the implications.

```{r}
df_h=fit5 %>% gather_draws(HIP[i]) %>% median_hdci(.width=.9) %>% select(i,.value)
df_j=goat_sum %>% select(-i_s,-i_e) %>% 
     left_join(df_h,by=c("farm_number"="i")) %>% rename(median_HIP=.value)

df_j %>% flextable()
```

### Farm Level Prevalance

We are also interested in understanding, _under the condition that a herd is infected, what is the prevalence of infection within that herd_ (CWHP).  Stated another way, what fraction of a herd is infected (when that number is not 0).

Recognizing that we have pretty sparse data for any particular herd (small sample size for the herd, small number of positives), the wide variability in the parameter estimate is not surprising.   However, we do see strong consistency if we focus on (median levels) especially if we look at herds that are likely infected.

If CWHP is relatively low, it means the disease is does not spread completely throught the herd.  *This phenomena needs to be better understood.*

```{r eval=FALSE, include=FALSE}
mcmc_areas(posterior,regex_pars=c("CWHP")) 
```

```{r eval=FALSE, include=FALSE}
fit5 %>% gather_draws(CWHP[i]) %>% median_hdi(.width=.9) %>% flextable()
```



## Is our model any good?

One way to address this question is to see if our model (with the fitted parameters) produces data "like" our original.  I will discuss this in more detail at our meeting.


```{r eval=FALSE, include=FALSE}
df_a=goat %>% count(elisa,pcr,culture) %>% arrange(elisa,pcr,culture)
df_a %>% flextable()
```

Let's create a helper function that takes a sampled parameter draw and generates a replicate sample:

```{r eval=FALSE, include=FALSE}
this_samp=2222
make_rep <- function(this_samp) {
  A=posterior[this_samp,]
  
  pi0=rbernoulli(27,A["HTP"])
  
  cwhp=A[1:27]
  spec=A[c("spec[1]","spec[2]","spec[3]")]
  sens=A[c("sens[1]","sens[2]","sens[3]")]
  
  #pi1=(pi0) %*% t(1-spec) 
  
  
  
  y_rep=matrix(,nrow=316,ncol=3)
    for (n in 1:27) {
  
      for (k in data5$I_s[n]:data5$I_e[n]) {
           # is this sheep infected determines value for pi1
           if (!pi0[n]) {  # herd not infected
                pi1=1-spec
           } else if (rbernoulli(1,cwhp[n])) { #sheep infected
                pi1=sens
           } else {
               pi1=1-spec 
           }
           
           
           
           y_rep[k,1:3]=rbernoulli(3,pi1)
           }
      }
  
  y_rep=as.tibble(y_rep+0) %>% rename(elisa=V1,pcr=V2,culture=V3)
}
```

```{r eval=FALSE, include=FALSE}

make_rep(sample(5000,1)) %>%  count(elisa,pcr,culture) %>% arrange(elisa,pcr,culture) %>% flextable()
```


# Grant deliverables

### Visualizing the data

Let's first create a visual of the entire dataset, illustrating sparcity of positive samples.

```{r echo=FALSE, fig.cap="Tests taken on male goats is bosed in black."}

goat_r=goat_s %>%
     mutate(elisa=if_else(valid_elisa==1,as.logical(elisa),NA),
            pcr=if_else(valid_pcr==1,as.logical(pcr),NA),
            culture=if_else(valid_culture==1,as.logical(culture),NA))

g2=goat_r %>% group_by(farm_number) %>% mutate(goat_num=row_number()) %>%
     select(-sample_number) %>% pivot_longer(elisa:culture, names_to = "test",values_to = "result" ) %>%
     mutate(result=as.factor(result)) %>% 
     left_join(df_j %>% select(farm_number,median_HIP))




CPCOLS <- c("#FFEFDB", "#5F9EA0","#D1D1D1")
CPCOLS2 <- c("#D1D1D1", "#000000")
p1g=g2 %>% ggplot(aes(x=goat_num,y=fct_inorder(test),fill=result)) +
     facet_wrap(~fct_reorder(as.factor(farm_number),median_HIP,.desc=TRUE),
                ncol=4,strip.position="left",dir="v")+
     geom_tile(aes(color=sex), show.legend = FALSE,width=.9)+
     scale_fill_manual(values = CPCOLS)+
     scale_color_manual(values=CPCOLS2)+
     theme_minimal() +
     theme(strip.background = element_blank(),
           strip.placement = "outside",
           strip.text.y.left = element_text(size = 8,angle = 0,color="gray72"),
           axis.title.y=element_text(color="gray72")) +
     xlab("number of goats tested on farm")+ylab("Farm Number")+
     ggtitle("Results of MAP testing of goat herds on 27 farms in grant study area")

p1g


#ggsave(filename=here("figs","data_goat.png"), width=12,height=8,units="cm"   )
     
```

```{r eval=FALSE, include=FALSE}
g2_cross=g2 %>% tabyl(sex,result) %>% rename(neg="FALSE",pos="TRUE") %>%
     adorn_percentages("row") %>% 
  adorn_pct_formatting(rounding = "half up", digits = 0) %>%
  adorn_ns() 
g2_cross %>% flextable()
```

### Bayesian Analysis

In Bayesian statistics, the key idea is that for unknown quantities, the analysis tries to capture the *uncertainty* not by giving a single value for those unknown quantities, but by showing a distribution (or histogram plot) of values that arise from repeated simutlation.   The values that result more often in the simulations are plotted with a higher likelilihood.  By plotting the whole distribution, we get an better understanding of not only what are the likely values, but what is the range of reasonable values.  Key quantities from that analysis are plotted below.

### Herd True Prevelence

Herd True Prevelence (HTP) describes the fraction of herds that have at least one case of MAP infection/Johnes in the herd.   HTP could be directly measured if:

* Testing were perfect.
* All farms were tested.
* All goats on the farm were tested.

This parameter is estimated from the data based on a standard model.

```{r echo=FALSE}
plot_title <- ggtitle("Herd Level Prevalence",
                      "with medians and 80% intervals")
mcmc_areas(posterior,prob=.8,pars="HTP") + plot_title

```
As summary information, we provide a point estimate (median) along with an 90% hdi interval (highest density interval) estimate.   The median value indicates that about 30% of the farms are infected.
```{r echo=FALSE}

fit5 %>% gather_draws(HTP) %>% median_hdi(.width=.9) %>% flextable()
```

### Herd Infection Probability (HIP)

For each "herd" (farm), we may estimate the probability that the farm has an infected animal (within the group of animals being tested).  We call this value *Herd Infection Probability* (HIP) and we estimate its value separately for each farm.  As a Bayesion parameter, it is also shown as a distribution (of values).  The value for HTP reflects an overall regional average, while HIP reflects the conditions on a farm and is a result or the data collected from that farm, as adjusted for the understanding of the accuracy of the tests and the overall conditions expected in the region.

```{r echo=FALSE}
farm_list=c(9,2,4,5,10,3,1,28)
plot_title <- ggtitle("HIP distribution")
p2g1=mcmc_areas(posterior,pars=vars(param_range("HIP",farm_list)))+plot_title
p2g2=g2 %>%  filter(farm_number %in% farm_list) %>%  ggplot(aes(x=goat_num,y=fct_inorder(test),fill=result)) +
     facet_wrap(~fct_reorder(as.factor(farm_number),median_HIP,.desc=TRUE),
                ncol=1,strip.position="left",dir="v")+
     geom_tile(color="grey", show.legend = FALSE)+
     scale_fill_manual(values = CPCOLS)+
     theme_minimal() +
     theme(strip.background = element_blank(),
           strip.placement = "outside",
           strip.text.y.left = element_text(size = 9,angle = 0,color="gray72"),
           axis.title.y=element_text(color="gray72")) +
     xlab("")+ylab("Farm Number")+
     ggtitle("Results of MAP testing")
p2g1+p2g2+ 
  plot_layout(widths = c(1.5, 1))


#ggsave(filename=here("figs","HIP_goat.png"), width=8,height=6,units="cm"   )
#ggsave(filename=here("figs","HIP_goat.tiff"), width=8,height=6,units="cm" ,dpi=300  )

```
As explanation of above figure: On the left is the HIP distribution for a selection of farms, and on the right is the associated testing data, where we have chosen farms to illustrate some key ideas:

* On farm 9, the positive tests were all on the same sheep which essential confirms the presence of MAP, so the infection probability is higher than on farm 2, even though it has the same number and type of positive tests.
* Comparing farm 4 and farm 10, we see the single postive ELISA pushes the HIP probability distribution (and mean value) toward higher probabilities of infection.
* For farms 5, 10, and 3, note that as we increase the number of goats testes, the HIP shifts to the left, and the median probability is reduced.
* Note that overall, even though no positive tests were seen on farm three, the HIP is not 0.  (The medial value on farm 3 is actually 10%).  This risk of undetected MAP persists due to the known low sensitivity of the testing procedures in an operational setting.

### Farm 28 shows extremely low liklihood of infection.

```{r echo=FALSE}
farm_list=c(28)
plot_title <- ggtitle("HIP distribution")
p2g1=mcmc_areas(posterior,pars=vars(param_range("HIP",farm_list)))+plot_title
p2g2=g2 %>%  filter(farm_number %in% farm_list) %>%  ggplot(aes(x=goat_num,y=fct_inorder(test),fill=result)) +
     facet_wrap(~fct_reorder(as.factor(farm_number),median_HIP,.desc=TRUE),
                ncol=1,strip.position="left",dir="v")+
     geom_tile(color="grey", show.legend = FALSE)+
     scale_fill_manual(values = CPCOLS)+
     theme_minimal() +
     theme(strip.background = element_blank(),
           strip.placement = "outside",
           strip.text.y.left = element_text(size = 12,angle = 0,color="gray72"),
           axis.title.y=element_text(color="gray72")) +
     xlab("number of goats tested on farm")+ylab("Farm Number")+
     ggtitle("Results of MAP testing")
p2g1+p2g2+ 
  plot_layout(widths = c(1.5, 1))
```



### HIP median values for full dataset

We re-vizualize the collected testing data, annotating with the median value of herd infection probability.

```{r echo=FALSE}
p1g+
     geom_text(aes(x=15,y="pcr",label=formattable::digits(median_HIP,2))) +
     ggtitle("MAP Testing results (annonated with median HIP values)")
```

### (Conditional) Within Herd Prevalence

Whereas *herd prevelence* (HTP) is concerned with what fraction of herds will have an infected animal, the other prevelence measure of interest is *what fraction of goats are carrying the infection*?  In the typical model, the parameter of interest considers this questions on a *per farm* basis, and attempts to explain ... under the assumption that MAP is present in the herd, what fraction of the herd is infected?   This quantity is designated as the *conditional within herd prevalence* (CWHP) and is estimate, as a distribution, herd by herd for each farm, again, based on the testing outcomes.

```{r}
farm_list=c(22,24,15,8,11,1,28)
plot_title <- ggtitle("CWHP distribution")
p3g1=mcmc_areas(posterior,pars=vars(param_range("CWHP",farm_list)))+plot_title+ geom_vline(xintercept = .25)
p3g2=g2 %>%  filter(farm_number %in% farm_list) %>%  ggplot(aes(x=goat_num,y=fct_inorder(test),fill=result)) +
     facet_wrap(~fct_reorder(as.factor(farm_number),median_HIP,.desc=TRUE),
                ncol=1,strip.position="left",dir="v")+
     geom_tile(color="grey", show.legend = FALSE)+
     scale_fill_manual(values = CPCOLS)+
     theme_minimal() +
     theme(strip.background = element_blank(),
           strip.placement = "outside",
           strip.text.y.left = element_text(size = 12,angle = 0,color="gray72"),
           axis.title.y=element_text(color="gray72")) +
     xlab("number of goats tested on farm")+ylab("Farm Number")+
     ggtitle("Results of MAP testing")
p3g1+p3g2+ 
  plot_layout(widths = c(1.5, 1))
```

### Median WHIP for full dataset


```{r}
#df_h=fit5 %>% gather_draws(HIP[i]) %>% median_hdci(.width=.9) %>% 

df_hc=fit5 %>% gather_draws(CWHP[i]) %>% median_hdci(.width=.9) %>% select(i,.value)


df_jc=goat_sum %>% select(-i_s,-i_e) %>% 
     left_join(df_hc,by=c("farm_number"="i")) %>% rename(median_CWHP=.value)

g3=goat_s %>% group_by(farm_number) %>% mutate(goat_num=row_number()) %>%
     select(-sample_number) %>% pivot_longer(elisa:culture, names_to = "test",values_to = "result" ) %>%
     mutate(result=as.logical(result)) %>% left_join(df_jc %>% select(farm_number,median_CWHP))
g3 %>%    ggplot(aes(x=goat_num,y=fct_inorder(test),fill=result)) +
     facet_wrap(~fct_reorder(as.factor(farm_number),median_CWHP,.desc=TRUE),
                ncol=4,strip.position="left",dir="v")+
     geom_tile(color="grey", show.legend = FALSE)+
     scale_fill_manual(values = CPCOLS)+
     theme_minimal() +
     theme(strip.background = element_blank(),
           strip.placement = "outside",
           strip.text.y.left = element_text(size = 12,angle = 0,color="gray72"),
           axis.title.y=element_text(color="gray72")) +
     xlab("number of goats tested on farm")+ylab("Farm Number")+
     ggtitle("Results of MAP testing")+
     geom_text(aes(x=15,y="pcr",label=formattable::digits(median_CWHP,2))) +
     ggtitle("MAP Testing results (annonated with median CWHP values)")

```


### Age and Sex 
