---
title: "Using stan with heirarchical model"
author: "Joe Skufca"
date:  "2021-12-23"
output: 
    html_notebook:
       toc: true
---

This experiment intends to analyze data from each farm as a herd, with common specificity and sensititive.  We assume on a single test (for which I will use the culture data)    The dataset of focus here is the "goat" dataset.

### Setup

```{r setup}
options(digits = 3)
knitr::opts_chunk$set( fig.width = 6,highlight=T)

library(tidyverse)
library(janitor)
library(PriorGen)


library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores(logical = FALSE))

library(bayesplot)
library(here)
library(tidybayes)
```

## Example 1

We will analyze the goat data, but without a heirarchical model (so no need to include which farm).

### Data

Load data.  We eliminate the one sample that was questionable.


```{r}
goat_file=here("data","goat_results.xlsx")

goat = readxl::read_excel(goat_file, sheet = "Sheet1") %>%
     clean_names() %>%
     select(sample_number,farm_number,  elisa, pcr, culture) %>%
     filter(elisa  %in% c("p", "n")) %>%
     filter(culture %in% c("p", "n")) %>%
     mutate(across(elisa:culture, ~ (.x == "p") + 0)) 


```

### Assemble Stan model data

In addition to the data, we need information for prior assumptions regarding sensitivity and specificity.

```{r}
goat_sum=goat %>% rowid_to_column("id") %>% group_by(farm_number) %>%
     summarise(i_s=first(id),i_e=last(id))
```




```{r}
# prior from Ontario 2011 paper
data5= list(F=27,
            I=316,
            I_s=goat_sum$i_s,
            I_e=goat_sum$i_e,
            J=3,
            y=as.matrix(goat %>% select(3:5)),
            alpha_sens = c( 4.33,   2.41,    4.98),
            beta_sens =  c( 8.77,   4.29,    6.96),
            alpha_spec = c(99.7 , 107.2 , 2291.17),
            beta_spec=   c( 6.19,   3.17,    6.74),
            htp_alpha=1.55,
            htp_beta=2.66,
            b1_mu=-2,
            b1_sd=2,
            eta_mu=0,
            eta_sd=10
                    )

  # int<lower=1> F; //number of farms
  # int<lower=1> I; //total number of animals tested
  # int<lower=1> I_s[F]; //starting index for animals from from f
  # int<lower=1> I_e[F]; //ending index for animals from from f
  # 
  # int<lower=1> J; // number of tests - 3 for our work
  # int<lower=1,upper=2> y[I,J]; // test resuls  0 for neg, 1 for pos
  # // params for prior for sensitivities and specificities
  # row_vector<lower=0>[J] alpha_sens;
  # row_vector<lower=0>[J] beta_sens;
  # row_vector<lower=0>[J] alpha_spec;
  # row_vector<lower=0>[J] beta_spec;
```


### fitting the model

```{r}
fit5=stan(file=here("src","heir_multi_test.stan"),data=data5,
          warmup=500,chains=12,iter=2000)
```

###  Visualizing results

```{r}
posterior=as.matrix(fit5)
```


Prevalence

```{r}
plot_title <- ggtitle("Posterior distributions",
                      "with medians and 80% intervals")
mcmc_areas(posterior,prob=.8,pars="HTP")


```

```{r}
summary(fit5)
mcmc_areas(posterior,prob=.8,pars="beta1")
```

### Which farms?

```{r}
mcmc_areas_ridges(posterior,regex_pars=c("EWHP\\[2[0-7]")) 
```
### Farm Level prevelence

```{r}
mcmc_areas_ridges(posterior,regex_pars=c("HID\\[1[0-5]"))
```
```{r}
mcmc_areas_ridges(posterior,regex_pars=c("sens"))
mcmc_areas_ridges(posterior,regex_pars=c("spec"))
```


### Chain mixing

```{r}
plot(fit5, plotfun = "trace", pars = c("beta1"), inc_warmup = TRUE)

```


```{r}

mcmc_intervals(posterior,regex_pars=c("eta"))

```



########################








