---
title: "Using stan with heirarchical model"
author: "Joe Skufca"
date:  "2021-12-23"
output: 
    html_notebook:
       toc: true
---

Experiment 6 uses re-parameterization of between herd prevalence using a more standard approach (with beta distribution) from Liapi et al (2011). This experiment intends to analyze data from each farm as a herd, with common specificity and sensititive.  We assume on a single test (for which I will use the culture data)    The dataset of focus here is the "goat" dataset.

### Setup

```{r setup}
options(digits = 3)
knitr::opts_chunk$set( fig.width = 6,highlight=T)

library(tidyverse)
library(janitor)
library(PriorGen)


library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores(logical = FALSE))

library(bayesplot)
library(here)
library(tidybayes)
library(shinystan)
```

## Example 1

We will analyze the goat data, but without a heirarchical model (so no need to include which farm).

### Data

Load data.  We eliminate the one sample that was questionable.


```{r}
goat_file=here("data","goat_results.xlsx")

goat = readxl::read_excel(goat_file, sheet = "Sheet1") %>%
     clean_names() %>%
     select(sample_number,farm_number,  elisa, pcr, culture) %>%
     filter(elisa  %in% c("p", "n")) %>%
     filter(culture %in% c("p", "n")) %>%
     mutate(across(elisa:culture, ~ (.x == "p") + 0)) 


```

### Assemble Stan model data

In addition to the data, we need information for prior assumptions regarding sensitivity and specificity.

```{r}
goat_sum=goat %>% rowid_to_column("id") %>% group_by(farm_number) %>%
     summarise(i_s=first(id),i_e=last(id))
```




```{r}
# prior from Ontario 2011 paper
data5= list(F=27,
            I=316,
            I_s=goat_sum$i_s,
            I_e=goat_sum$i_e,
            J=3,
            y=as.matrix(goat %>% select(3:5)),
            alpha_sens = c( 4.33,   2.41,    4.98),
            beta_sens =  c( 8.77,   4.29,    6.96),
            alpha_spec = c(99.7 , 107.2 , 2291.17),
            beta_spec=   c( 6.19,   3.17,    6.74),
            htp_alpha=1.55,
            htp_beta=2.66,
            mu_alpha=2.15,
            mu_beta=7.52,
            psi_a=5.4,
            psi_b=.55
                    )

  # int<lower=1> F; //number of farms
  # int<lower=1> I; //total number of animals tested
  # int<lower=1> I_s[F]; //starting index for animals from from f
  # int<lower=1> I_e[F]; //ending index for animals from from f
  # 
  # int<lower=1> J; // number of tests - 3 for our work
  # int<lower=0,upper=1> y[I,J]; // test resuls  0 for neg, 1 for pos
  # // params for prior for sensitivities and specificities
  # row_vector<lower=0>[J] alpha_sens;
  # row_vector<lower=0>[J] beta_sens;
  # row_vector<lower=0>[J] alpha_spec;
  # row_vector<lower=0>[J] beta_spec;
  # real htp_alpha; // prior  (HTP is tau from Liapi)
  # real htp_beta; // prior
  # real mu_alpha;// prior
  # real mu_beta;// prior
  # real psi_a;// prior
  # real psi_b;// prior
```


### fitting the model

```{r}
fit5=stan(file=here("src","heir_multi_test_B.stan"),data=data5,
          warmup=500,chains=12,iter=2000)
```

###  Visualizing results

```{r}
posterior=as.matrix(fit5)
```


Prevalence

```{r}
plot_title <- ggtitle("Posterior distributions",
                      "with medians and 80% intervals")
mcmc_areas(posterior,prob=.8,pars="HTP")


```

```{r}
summary(fit5)
#mcmc_areas(posterior,prob=.8,pars="beta1")
```

### Which farms?

```{r}
mcmc_areas_ridges(posterior,regex_pars=c("CWHP")) 
```
### Farm Level prevelence

```{r}
mcmc_areas_ridges(posterior,regex_pars=c("HID\\[[1-9],1"))
```
```{r}
mcmc_areas_ridges(posterior,regex_pars=c("sens"))
mcmc_areas_ridges(posterior,regex_pars=c("spec"))
```


### Chain mixing

```{r}
plot(fit5, plotfun = "trace", pars = c("HTP"), inc_warmup = TRUE)

```





## Understanding our data

Let's look at a table of our data.  We will want to see if our model can generate data "like" ours, so ... we need to be able to characterize.

We will use gtsummary package to assist.
```{r}
library(gtsummary)
```

```{r}
df_a=goat %>% count(elisa,pcr,culture) %>% arrange(elisa,pcr,culture)
df_a
```

Let's create a helper function that takes a sampled parameter draw and generates a replicate sample:

```{r}
this_samp=2222
make_rep <- function(this_samp) {
  A=posterior[this_samp,]
  
  pi0=rbernoulli(27,A["HTP"])
  
  cwhp=A[1:27]
  spec=A[c("spec[1]","spec[2]","spec[3]")]
  sens=A[c("sens[1]","sens[2]","sens[3]")]
  
  #pi1=(pi0) %*% t(1-spec) 
  
  
  
  y_rep=matrix(,nrow=316,ncol=3)
    for (n in 1:27) {
  
      for (k in data5$I_s[n]:data5$I_e[n]) {
           # is this sheep infected determines value for pi1
           if (!pi0[n]) {  # herd not infected
                pi1=1-spec
           } else if (rbernoulli(1,cwhp[n])) { #sheep infected
                pi1=sens
           } else {
               pi1=1-spec 
           }
           
           
           
           y_rep[k,1:3]=rbernoulli(3,pi1)
           }
      }
  
  y_rep=as.tibble(y_rep+0) %>% rename(elisa=V1,pcr=V2,culture=V3)
}
```

```{r}

make_rep(sample(5000,1)) %>%  count(elisa,pcr,culture) %>% arrange(elisa,pcr,culture)
```





