---
title: "Exp 17:Using stan with heirarchical model"
author: "Joe Skufca"
date: "2025-06-09"
output:
  html_notebook:
    toc: true
  word_document:
    toc: true
---


Experiment 17 examines the collaborative dataset working with the Wisconsin lab.

We analyze using 4 tests (Culture, PCR, and 2 Elisa tests).

Starting from the collected data from the grant funded study, we have removed animals for 
which a complete set of tests results was not available.

>The data is from file 
>"MAP NY Goat Project 2021-jc-2025 - WI results - complete for analysis.xlsx"
> dtd 2025-06-08  
> as copied to file "goat_results2_wi.xlsx"


## Analytic approach

* We use a two-latent class model, which assumes each animal is classes as either "infected" or "non-infected."
* We make estimates of within herd prevalence for each herd.
* Prior for herd level prevalence was is described below and is based on lit review.  
* Prior for test sensitivity and specificity are support by literature review and are described below.



## Setup

We load anticipated packages and set rstan options.

```{r setup}
options(digits = 3)
knitr::opts_chunk$set( highlight=T,message=FALSE, warning=FALSE,echo=FALSE)

library(tidyverse)
library(patchwork)
library(janitor)
library(PriorGen)
library(ggpmisc)

library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores(logical = FALSE))

library(bayesplot)
library(here)
library(tidybayes)
library(shinystan)
library(DT)
library(flextable)
```

## Preparing data for Bayesian Analysis

We will analyze the goat data, but using a hierarchical model (with each farm treated as a herd).

### Data

Load data.  


```{r message=FALSE, warning=FALSE, results="asis"}
goat_file=here("data","goat_results2_wi.xlsx")

goat= readxl::read_excel(goat_file) %>%
  clean_names() %>%
  tidyr::separate_wider_delim(cols=accession,
    delim="-",    names=c(NA,NA,"within_farm_number"),
    too_many = "drop"
    ) %>%
  mutate(within_farm_number=as.numeric(within_farm_number)) %>%
  filter(kit_a_elisa_s_p !="N/A") %>%
  mutate(elisa_a=case_when(
          str_starts(inter_kit_a,"N") ~ "n",
          TRUE ~ "p")) %>%
  mutate(elisa_b=case_when(
          str_starts(interpret_use_goat_kit_b,"N") ~ "n",
          TRUE ~ "p")) %>%
  rename(farm_number=farms) %>%
  mutate(pcr=case_when(
          str_starts(is900_pcr_res,"N") ~ "n",
          TRUE ~ "p")) %>%
  mutate(culture=case_when(
          str_starts(johnes_liquid_culture,"N") ~ "n",
          TRUE ~ "p")) %>%
  rowid_to_column("sample_number") %>%
  select(sample_number,farm_number,within_farm_number, elisa_a,elisa_b, pcr, culture) %>%
  mutate(across(elisa_a:culture, ~ (.x == "p") + 0)) %>%
  mutate(valid_elisa_a=1,valid_elisa_b=1,valid_pcr=1,valid_culture=1)


```





#### Farm level summary

I include here a breakdown table by farm, as it is part of assembling the data structure required to input to the stan code. This table will help us to understand why certain farms are deemed to be higher likelihood of being infected as well as differing levels of within farm prevalence.

> Farm numbers are as listed in the aggregated data file from Wisconsin.   That data "skips" numbers _19_ and _23_.
> Likely, we should fix before publication.

```{r}
goat_sum=goat %>% rowid_to_column("id") %>% group_by(farm_number) %>%
     summarise(i_s=first(id),i_e=last(id),n=n(),across(elisa_a:culture,sum)) %>%
     rowid_to_column("id")
goat_sum %>% flextable() 
```


### Development of (assumed) Bayesian "prior" information

Bayesian analysis requires that modeled parameters be described by an assumed _prior distribution_.
The prior can be used to capture problem understanding that exists before data collection.  
We use both literature review and expert local knowledge to develop the priors for this analysis.


We use the `PriorGen` package to assist with developing hyperparameters (the mathematical description of our prior distributions).  The wrapper functions created here allow output from that package to be used programmatically.

```{r }
my_findbeta=function(...){
     findbeta(...)$parameters
}

my_findbetamupsi=function(...){
     mupsi_obj=findbetamupsi(...)
     c(mupsi_obj$param_beta,mupsi_obj$param_gamma)
}

```

#### Hyperparameter for priors

For several of the hyperparameters, we will use Beta or Gamma distributions as priors (the typical approach for the literature for this type of latent class analysis), using expert knowledge to develop informative priors.  

##### Herd True Prevelance (HTP): Prior


```{r}
htp1=.3
htp2=.67

```
 
For HTP hyperparameters: Based on expert opinion, we suspect that the most likely value for the fraction of farms with Johnes present in the goat flock is about `r formattable::percent(htp1)`, and we are **95%** certain that the fraction of infected farms is less that `r formattable::percent(htp2)`.  We model the prior distribution using a beta function, shown below. 
 
```{r fig.width=3}
htp_hyp=my_findbeta(themode = htp1,percentile.value = htp2,lower.v = TRUE)

#htp_hyp=findbeta(themode = htp1,percentile.value = htp2,lower.v = TRUE)$parameters
ggfortify::ggdistribution(dbeta,seq(0,1,.005),shape1=htp_hyp[1],shape2=htp_hyp[2])+
     theme_minimal()+
     ggtitle("Assume prior for herd level prevalence")
```


##### Within Herd Prevalence (WHP): Prior

Within herd prevalence prior information is captured by both a mean value and a description of the variance, with hyperparameters $\mu$ and $\psi$ drawn from a Beta and Gamma distribution respectivelvy. 
To generate the hyperparameters for $\mu$ and $\psi$ so that we can prescribe priors for within herd prevalence, we complete the  portions of this sentence with expert opinion based on area knoweledge:

```{r}
v1=.2 
v2=.95
v3= .33
v4=.9
v5=.5
v6=.7

mupsi_hyp=my_findbetamupsi(themean=v1, percentile=v2, percentile.value=v3,
 lower.v=TRUE,
psi.percentile=v4, percentile.median=v5, percentile95value=v6)


```

> Consider herds in this region of study, and consider those farms for which it is known that
> Johnes is present in that herd.  Focusing on the __fraction__ of animals on that farm that will be infected:
> the **average** value is thought to be `r formattable::percent(v1)`  and we are 
> `r formattable::percent(v2)` confident that **average value** it is not more than `r formattable::percent(v3)`.
> Rather than the *average value*, we may estimate an upper bound, such that 
> `r formattable::percent(v4)` of the infected farms would probably have an infection rate below
> `r formattable::percent(v5)`, and we are 95% certain that they would not exceed `r formattable::percent(v6)`.

##### Sensitivity and Specificity: (Prior)

We also need prior for sensitivity and specificity for each of the tests. 

> __Background.__  One of the challenges here is that we are nominally using the test as measures of the presence of infection.  The tests used here have a good bit of independence (using different biological signals to produce a test outcome) which is an assumption of the statistical model that is used.  However, the infection progression in the body is such that not all of the signals would necessarily be present at the same time, such that the assumption that a "diseased" goat should test positive on all test is not a precise description of the biology.  As such, the effect (in an operational setting) is that the test would typically show lower accuracy than they would compared to the "gold standard" for that particular type of test.  (For example: The manufacturer's statements about accuracy for PCR are in comparison to a gold standard PCR test - equivalently does the PCR test being used detect DNA material when bacterial DNA material is actually present in the blood at a level above some known minimum. But the biological situation may be that the host is shedding bacteria in the fecal matter at a level which can be cultured even when not detectable in the blood, as this shedding results from a different biological process.)



To allow for this reduced accuracy within our Bayesian analysis, we widen our assumed prior distributions, making them less informative.  We use the modes as suggested, but using 95% intervals that allow for a much broader range of error, allowing the data to have a stronger influence on the posterior distribution.


For each of these parameters, we use a template that follows this exemplar pattern:

For parameter xxx - use mode of (   ), with 95% probability that is value is above (   )%.


```{r}
# for each test, the first two values relate to sensitivity, the second to specificy
elisa_v=c(1,.33,.985,.9)
pcr_v=c(.956,.33,1,.9) 
culture_v=c(.231,.1,.99,.9) 

elisa_sens=my_findbeta(themode = elisa_v[1],percentile.value = elisa_v[2],lower.v = FALSE)
elisa_spec=my_findbeta(themode = elisa_v[3],percentile.value = elisa_v[4],lower.v = FALSE)
pcr_sens=my_findbeta(themode = pcr_v[1],percentile.value = pcr_v[2],lower.v = FALSE)
pcr_spec=my_findbeta(themode = pcr_v[3],percentile.value = pcr_v[4],lower.v = FALSE)
culture_sens=my_findbeta(themode = culture_v[1],percentile.value = culture_v[2],lower.v = FALSE)
culture_spec=my_findbeta(themode = culture_v[3],percentile.value = culture_v[4],lower.v = FALSE)


elisa_a_sens=elisa_sens
elisa_b_sens=elisa_sens
elisa_a_spec = elisa_spec
elisa_b_spec = elisa_spec

```

* Elisa sens: use mode of `r formattable::percent(elisa_v[1])`, with 95% probability that is value is above `r formattable::percent(elisa_v[2])`
* Elisa spec: use mode of `r formattable::percent(elisa_v[3])`, with 95% probability that is value is above `r formattable::percent(elisa_v[4])`
* pcr sens: use mode of `r formattable::percent(pcr_v[1])`, with 95% probability that is value is above `r formattable::percent(pcr_v[2])`
* pcr spec: use mode of `r formattable::percent(pcr_v[3])`, with 95% probability that is value is above `r formattable::percent(pcr_v[4])`
* culture sens: use mode of `r formattable::percent(culture_v[1])`, with 95% probability that is value is above `r formattable::percent(culture_v[2])`
* culture spec: use mode of `r formattable::percent(culture_v[3])`, with 95% probability that is value is above `r formattable::percent(culture_v[4])`

We assume the same prior for both ELISA tests.

### Assemble data for Stan model

We now assemble the data for the Stan model.  The data structure is described in the Stan code, but we will summarize here.

```{r}

data5= list(F=nrow(goat_sum), # number of farms
            I=nrow(goat), # total number of animals tested
            I_s=goat_sum$i_s, # starting index for animals from from f
            I_e=goat_sum$i_e, # ending index for animals from from f
            J=4, # number of tests - 4 for our work
            y=as.matrix(goat %>% select(elisa_a:culture)), # test results  0 for neg, 1 for pos
            v=as.matrix(goat %>% select(valid_elisa_a:valid_culture)), 
            # validity of test results, allows removal of invalid tests
            # allowing that not all tests performed on all animals
            alpha_sens = c( elisa_a_sens[1], elisa_b_sens[1], # hyperparameters for sensitivity
                            pcr_sens[1],    culture_sens[1]),
            beta_sens =  c( elisa_a_sens[2], elisa_b_sens[2], 
                            pcr_sens[2],    culture_sens[2]),
            alpha_spec = c( elisa_a_spec[1], elisa_b_spec[1], # hyperparameters for specificity
                            pcr_spec[1],    culture_spec[1]),
            beta_spec=   c( elisa_a_spec[2], elisa_b_spec[2],
                            pcr_spec[2],    culture_spec[2]),
            htp_alpha=htp_hyp[1], # hyperparameters for herd true prevalence
            htp_beta=htp_hyp[2],
            mu_alpha=mupsi_hyp[1], # hyperparameters for within herd prevalence
            mu_beta=mupsi_hyp[2],
            psi_a=mupsi_hyp[3],
            psi_b=mupsi_hyp[4]
                    )

```





### fitting the model

To generate robust statistics, we generate multiple chains, each generating many samples.  
The analytic methods will average over these collective statistics. 
```{r}
system.time({
fit5=stan(file=here("src","heir_multi_test_B_OOS_2.stan"),data=data5,
          warmup=500,chains=12,iter=4000)})

```


```{r}
df_h=fit5 %>% gather_draws(HIP[i]) %>% median_hdci(.width=.9) # %>% select(i,.value)
df_j=goat_sum %>% select(-i_s,-i_e) %>% 
     bind_cols(df_h) %>% rename(median_HIP=.value)
```



##  Analysis Results

Let's first visualize of the analyzed data set.  It is important to
acknowledge the sparsity of positive samples.

```{r echo=FALSE, fig.cap="Goat testing results (analyzed dataset)"}
goat_r=goat%>%
     mutate(elisa_a=if_else(valid_elisa_a==1,as.logical(elisa_a),NA),
            elisa_b=if_else(valid_elisa_b==1,as.logical(elisa_b),NA),
            pcr=if_else(valid_pcr==1,as.logical(pcr),NA),
            culture=if_else(valid_culture==1,as.logical(culture),NA))

g2=goat_r %>% group_by(farm_number) %>% mutate(goat_num=row_number()) %>%
     select(-sample_number,-within_farm_number) %>% 
     pivot_longer(elisa_a:culture, names_to = "test",values_to = "result" ) %>%
     mutate(result=as.factor(result)) %>% 
     left_join(df_j %>% select(farm_number,median_HIP))

CPCOLS <- c("#FFEFDB", "#5F9EA0","#D1D1D1")
CPCOLS2 <- c("#D1D1D1", "#000000")
p1g=g2 %>% ggplot(aes(x=goat_num,y=fct_inorder(test),fill=result)) +
     facet_wrap(~fct_reorder(as.factor(farm_number),median_HIP,.desc=TRUE),
                ncol=4,strip.position="left",dir="v")+
     geom_tile(aes(), show.legend = FALSE,width=.9)+
     scale_fill_manual(values = CPCOLS)+
     scale_color_manual(values=CPCOLS2)+
     theme_minimal() +
     theme(strip.background = element_blank(),
           strip.placement = "outside",
           strip.text.y.left = element_text(size = 8,angle = 0,color="gray72"),
           axis.title.y=element_text(color="gray72")) +
     xlab("number of goats tested on farm")+ylab("Farm Number")+
     ggtitle("Results of MAP testing of goat herds on 27 farms in grant study area")

p1g


#ggsave(filename=here("figs","data_goat.png"), width=12,height=8,units="cm"   )
     
```

> In Bayesian statistics, the key idea is that for unknown quantities, the analysis tries to capture the *uncertainty* not by giving a single value for those unknown quantities, but by showing a distribution (or histogram plot) of values that arise from repeated simulation.   The values that result more often in the simulations are plotted with a higher likelihood.  By plotting the whole distribution, we get an better understanding of not only what are the likely values, but what is the range of reasonable values.  Key quantities from that analysis are plotted below, along with tabluar information, where appropriate.




Primarily, here, I will be showing results with graphs, using a bit of words to describe.  

The plots will all be posterior distribution of various parameter values.

We start by looking simply at the implied capability of the tests themselves
(sensitivity and specificity), though we recognize that these statistics give are
commenting upon the "in-the-wild" performance of these tests from the farm perspective,
and are a different viewpoint that the laboratory specifications.

```{r}
posterior=as.matrix(fit5)
```




### Sensitivity and Specificity (Posterior distributions)

For our test set, we note that the small number of positive test in the sample imply
that assessment of these parameters will likely not be overly precise.  
Bayesian analysis is meant to hightlight the uncertainty in these estimates, and we will see that
the posterior distributions are quite wide, especially for sensitivity.


#### Sensitivity: Do tests find the animals that have the disease?

Previous studies have identified that the sensitivity of these test is somewhat lower than desirable, 
with a reasonably large number of false negatives, implying that a single test may not
"clear" an animal.

Our results are consistent with those previous studies.

```{r}
test_set_labels = c("elisa_a","elisa_b","pcr","cult")
sensitivity_labels <- paste0(test_set_labels, "_sens")

plot_title <- ggtitle("Sensitivy (Goats)")
mcmc_areas(posterior,regex_pars=c("sens"))+
     scale_y_discrete(labels=sensitivity_labels)+plot_title

#ggsave(filename=here("figs","sens_goat.png"), width=7.5,height=5,units="cm"   )
```

#### Specificity: Can we avoid false positives?



Specificity is very high.

```{r }



plot_title <- ggtitle("Specificity (Goats)")
#ggsave(filename=here("figs","spec_goat.png"), width=10,height=8,units="cm"   )
mcmc_areas(posterior,regex_pars=c("spec"))+
     scale_y_discrete(labels= paste0(test_set_labels, "_spec"))+plot_title


#ggsave(filename=here("figs","spec_goat.png"), width=7.5,height=5,units="cm"   )
```



#### Tabular summary of sensitivy and specifity

```{r}
fit5 %>%
  gather_draws(spec[i], sens[i]) %>%
  median_hdci(.width = .9) %>%
  arrange(.variable, i) %>%
  mutate(
    param = case_when(
      .variable == "sens" ~ paste0(test_set_labels[i], "_sens"),
      .variable == "spec" ~ paste0(test_set_labels[i], "_spec")
    )
  ) %>%
  select(param, .value, .lower, .upper, .width) %>%
  flextable()
```




### Herd True Prevalence: posterior distribution

Herd True Prevalence (HTP) is the fraction of herds with at least one infected animal.

Herd True Prevalence (HTP) describes the fraction of herds that have at least one case of MAP infection/Johnes in the herd.   HTP could be directly measured if:

* Testing were perfect.
* All farms were tested.
* All goats on the farm were tested.

This parameter is estimated from the data based on a standard model.

```{r}
plot_title <- ggtitle("Herd Level Prevalence",
                      "with medians and 80% intervals")
mcmc_areas(posterior,prob=.8,pars="HTP") + plot_title
#ggsave(filename=here("figs","HTP_goat.png"), width=10,height=8,units="cm"   )
```
For a particular parameter, we can ask for detailed summary information in many different forms.  Below I provide a point estimate (median) along with an 90% hdci interval (highest density continuous interval) estimate.   The median value indicates that about 1/3 of the farms are infected.
```{r}

fit5 %>% gather_draws(HTP) %>% median_hdci(.width=.9) %>% flextable()
```

Note that the median value checks (somewhat) with the fraction of herds that had positive test results.





### Herd Infection Probability: posterior distribution

The model formulation allows computation of _herd infection probability_ (HIP), which gives the probability that a particular herd has at least one infected animal in the group of animals tested.  (Note the care of wording.  The model is not aware of the total number of animals on the farm, which would be necessary to more properly assess the full herd.)

For each "herd" (farm), we may estimate the probability that the farm has an infected animal (within the group of animals being tested).  We call this value *Herd Infection Probability* (HIP) and we estimate its value separately for each farm.  As a Bayesion parameter, it is also shown as a distribution (of values).  The value for HTP reflects an overall regional average, while HIP reflects the conditions on a farm and is a result or the data collected from that farm, as adjusted for the understanding of the accuracy of the tests and the overall conditions expected in the region.

```{r}
df_h=fit5 %>% gather_draws(HIP[i]) %>% median_hdci(.width=.9) # %>% select(i,.value)
df_j=goat_sum %>% select(-i_s,-i_e) %>% 
     bind_cols(df_h) %>% rename(median_HIP=.value)

dfhip = df_j %>% 
     arrange(-median_HIP) %>% 
     relocate(median_HIP) %>% 
     mutate(across(farm_number:culture, as.integer)) %>%
     select(-i) 

dfhip  %>% select(-(.width:.interval),-id,-.variable) %>%
     flextable()  %>%
     colformat_double(digits = 3) %>%
     vline(j=c(2,7)) %>%
     set_caption("Median Herd Infection Probability (HIP) with 90% interval")


```

```{r  fig.height=8, fig.width=4}
farm_list=dfhip$farm_number
farm_ids <- goat_sum %>%
  filter(farm_number %in% farm_list) %>%
  arrange(match(farm_number, farm_list)) %>%
  pull(id)
#farm_list=c(4,5,1,3)

# Build a named vector: names are the actual parameter names; values are the farm labels
param_names <- rev(paste0("HIP[", farm_ids, "]"))
farm_labels <- rev(paste0("Farm[", farm_list,"]"))

plot_title <- ggtitle("HIP distribution")
p_hip1 <- mcmc_areas(posterior,
                   pars=vars(param_range("HIP",rev(farm_ids))),
                   area_method = "scaled height") +
     scale_y_discrete(expand = expansion(add = c(0, 0)))
     plot_title 
     
p_res1=g2 %>%  filter(farm_number %in% farm_list) %>%  ggplot(aes(x=goat_num,y=fct_inorder(test),fill=result)) +
     facet_wrap(~fct_relevel(as.factor(farm_number),as.character(farm_list)),
                ncol=1,strip.position="left",dir="v")+
     geom_tile(color="grey", show.legend = FALSE)+
     scale_fill_manual(values = CPCOLS)+
     theme_minimal() +
     theme(strip.background = element_blank(),
           strip.placement = "outside",
           strip.text.y.left = element_text(size = 12,angle = 0,color="gray72"),
           axis.title.y=element_text(color="gray72")) +
     xlab("number of goats tested on farm")+ylab("Farm Number")+
     ggtitle("Results of MAP testing")

p_hip1+p_res1+ 
  plot_layout(widths = c(1.5, .8))

```

 
### Prevalance at the Farm Level (CWHP): posterior distribution

We are also interested in understanding, _under the condition that a herd is infected, what is the prevalence of infection within that herd_ (CWHP).  Stated another way, what fraction of a herd is infected if we *knew* that at least one animal was infected.

Recognizing that we have pretty sparse data for any particular herd (small sample size for the herd, small number of positives), the wide variability in the parameter estimate is not surprising.   However, we do see strong consistency if we focus on (median levels) especially if we look at herds that are likely infected.

```{r}
df_hp = fit5 %>% 
     gather_draws(CWHP[i]) %>% 
     median_hdci(.width=.9)

dfcwhp = goat_sum %>% 
         select(-i_s,-i_e) %>%
         bind_cols(df_hp) %>% rename(median_CWHP=.value) %>%
     arrange(-median_CWHP) %>%
     relocate(median_CWHP) %>%
     mutate(across(farm_number:culture, as.integer)) %>%
     select(-i)

dfcwhp %>% 
     select(-(.width:.interval),-id,-.variable) %>%
     flextable()  %>%
     colformat_double(digits = 3) %>%
     vline(j=c(2,7)) %>%
     set_caption("Median Conditional Within Herd Prevalence (CWHP) with 90% interval")
```




If CWHP is relatively low, it means the disease does not spread completely throughout the herd.  *This phenomena needs to be better understood.*

> Note: In the figure below, the similar shape on multiple farms is driven by the *assumption* that the disease is present. On farms with no postitive tests, the shape provides captures the situation that we have assumed that at least on animal is positive, but any  infected animals yielded a false negative.

```{r fig.width=5,fig.height=10}
# Use the same farm_list and farm_ids as above
# farm_list = dfhip$farm_number
# farm_ids <- ... (already defined above)

# Reverse the ids for matching vertical plotting order
param_names_cwhp <- rev(paste0("CWHP[", farm_ids, "]"))
farm_labels_cwhp <- rev(paste0("Farm[", farm_list, "]"))

plot_title_cwhp <- ggtitle("CWHP distribution")
p_cwhp1 <- mcmc_areas(
    posterior,
    pars = vars(param_range("CWHP", rev(farm_ids))),
    area_method = "scaled height"
  ) +
  scale_y_discrete(expand = expansion(add = c(0, 0))) +
  plot_title_cwhp+
  geom_vline(xintercept = .25) 

p_res1 <- g2 %>%
  filter(farm_number %in% farm_list) %>%
  ggplot(aes(x = goat_num, y = fct_inorder(test), fill = result)) +
  facet_wrap(
    ~fct_relevel(as.factor(farm_number), as.character(farm_list)),
    ncol = 1, strip.position = "left", dir = "v"
  ) +
  geom_tile(color = "grey", show.legend = FALSE) +
  scale_fill_manual(values = CPCOLS) +
  theme_minimal() +
  theme(
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text.y.left = element_text(size = 12, angle = 0, color = "gray72"),
    axis.title.y = element_text(color = "gray72")
  ) +
  xlab("number of goats tested on farm") + ylab("Farm Number") +
  ggtitle("Results of MAP testing")

p_cwhp1 + p_res1 + plot_layout(widths = c(1.5, .8))
```





## Additional figures 

The figures above are meant to be complete in capturing the analytic results.  
However, it may be useful to include some more focused data to illustrate particular
aspects of the result.

### Full dataset with HIP median labeling

We re-vizualize the collected testing data, annotating with the median value of herd infection probability.

```{r echo=FALSE}
p1g+
     geom_text(aes(x=15,y="pcr",label=formattable::digits(median_HIP,2))) +
     ggtitle("MAP Testing results (annonated with median HIP values)")
```






### Herd Infection Probability (HIP)

For each "herd" (farm), we may estimate the probability that the farm has an infected animal (within the group of animals being tested).  We call this value *Herd Infection Probability* (HIP) and we estimate its value separately for each farm.  As a Bayesion parameter, it is also shown as a distribution (of values).  The value for HTP reflects an overall regional average, while HIP reflects the conditions on a farm and is a result or the data collected from that farm, as adjusted for the understanding of the accuracy of the tests and the overall conditions expected in the region.


As explanation of above figure: On the left is the HIP distribution for a selection of farms, and on the right is the associated testing data, where we have chosen farms to illustrate some key ideas:

* On farm 9, the positive tests were all on the same sheep which essential confirms the presence of MAP, so the infection probability is higher than on farm 2, even though it has the same number and type of positive tests.
* Comparing farm 4 and farm 10, we see the single postive ELISA pushes the HIP probability distribution (and mean value) toward higher probabilities of infection.
* For farms 5, 10, and 3, note that as we increase the number of goats testes, the HIP shifts to the left, and the median probability is reduced.
* Note that overall, even though no positive tests were seen on farm three, the HIP is not 0.  (The medial value on farm 3 is actually 10%).  This risk of undetected MAP persists due to the known low sensitivity of the testing procedures in an operational setting.


```{r}
# Make sure farm_list2 and farm_ids2 are defined as before
farm_list2 <- c(4,17,11,9,5,3,7)
farm_order <- farm_list2   # c(4,17,11,9,5,3,7)

p_res2 <- g2 %>%
  filter(farm_number %in% farm_order) %>%
  ggplot(aes(x = goat_num, y = fct_inorder(test), fill = result)) +
  facet_wrap(
    ~fct_relevel(as.factor(farm_number), as.character(farm_order)),
    ncol = 1, strip.position = "left", dir = "v"
  ) +
  geom_tile(color = "grey", show.legend = FALSE) +
  scale_fill_manual(values = CPCOLS) +
  theme_minimal() +
  theme(
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text.y.left = element_text(size = 12, angle = 0, color = "gray72"),
    axis.title.y = element_text(color = "gray72")
  ) +
  xlab("number of goats tested on farm") + ylab("Farm Number") +
  ggtitle("Results of MAP testing")




# 50% equal-tailed intervals + medians for HIP[i] (plot only; leaves your 90% table untouched)
df_h50 <- fit5 %>%
  gather_draws(HIP[i]) %>%
  median_qi(.width = 0.50)  # 50% interval

df_j50 <- goat_sum %>%
  select(-i_s, -i_e) %>%      # same as you did earlier
  bind_cols(df_h50) %>%
  rename(median_HIP_50 = .value)

# Compact summary we’ll join to the density data

hip_summ50 <- df_j50 %>%
  transmute(
    i,
    parameter = paste0("HIP[", i, "]"),
    median    = median_HIP_50,
    lower50   = .lower,
    upper50   = .upper
  )


hip_farm_lookup <- tibble(
  farm_number = farm_order,
  parameter   = paste0("HIP[", farm_order, "]")
)

# density plot information
dens_df <- mcmc_areas_data(
  posterior,
  pars        = vars(param_range("HIP", farm_order)),
  prob_outer  = 1,
  area_method = "equal area"
  ) %>%
  left_join(hip_farm_lookup, by = "parameter") %>%
  mutate(    farm_number = factor(farm_number, levels = farm_order)) %>%
  left_join(hip_summ50, by = "parameter")

# median value for each farm
facet_df <- dens_df %>%
  distinct(farm_number, median)


# labeller function to show median value in facet labels
median_labeller <- function(farm_values) {
  med <- facet_df$median[match(farm_values, facet_df$farm_number)]
  sprintf("med = %.2f", med)
}



my_hip_plot <- ggplot(dens_df, aes(x = x, y = density)) +
     geom_line(color = "gray30", linewidth = 0.6) +
     geom_ribbon(
          data = ~ dplyr::filter(.x, x >= lower50 & x <= upper50),
          aes(ymin = 0, ymax = density),
          fill = "#4C78A8",
          alpha = 0.35,
          color = NA
     ) +
     geom_vline(
          data = dplyr::distinct(dens_df, farm_number, median),
          aes(xintercept = median),
          color = "black",
          linewidth = 0.6,
          linetype = "dashed"
     ) +
     facet_wrap(
          vars(farm_number),
          ncol = 1,
          scales = "free_y",
          strip.position = "left",
          labeller = labeller(farm_number = median_labeller)
     ) +
     coord_cartesian(xlim = c(0, 1)) +
     labs(
          title = "HIP distribution",
          subtitle = "Posterior with central 50% interval",
          x = "HIP",
          y = NULL
     ) +
     theme_minimal(base_size = 12) +
     theme(
          strip.text.y.left = element_text(
               size = 10,
               angle = 0,
               hjust = 1,
               color = "gray40"
          ),
          strip.background = element_blank(),
          
          axis.text.y  = element_blank(),
          axis.ticks.y = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor = element_blank()
     )



p_res2 + my_hip_plot + plot_layout(widths = c(.8, 1.2))

#ggsave(filename=here("figs","hip_sel_goat.png"), width=8,height=6,units="cm"   )
#ggsave(filename=here("figs","hip_sel_goat.tiff"), width=8,height=6,units="cm" ,dpi=300  )

```




### (Conditional) Within Herd Prevalence

Whereas *herd prevalence* (HTP) is concerned with what fraction of herds will have an infected animal, the other prevalence measure of interest is *what fraction of goats are carrying the infection*?  In the typical model, the parameter of interest considers this questions on a *per farm* basis, and attempts to explain ... under the assumption that MAP is present in the herd, what fraction of the herd is infected?   This quantity is designated as the *conditional within herd prevalence* (CWHP) and is estimate, as a distribution, herd by herd for each farm, again, based on the testing outcomes.

```{r echo=FALSE}
# Define your limited farm list
# sorted by CWHP 
# farm_list2 <- c(4,9,17,11,5,3,7)  # <- Use your desired farm numbers

# sorted same as HIP
farm_list2 <- c(4,17,11,9,5,3,7)


farm_order <- farm_list2

# Associated tile plot for limited farms
p_res2 <- g2 %>%
  filter(farm_number %in% farm_list2) %>%
  ggplot(aes(x = goat_num, y = fct_inorder(test), fill = result)) +
  facet_wrap(
    ~fct_relevel(as.factor(farm_number), as.character(farm_list2)),
    ncol = 1, strip.position = "left", dir = "v"
  ) +
  geom_tile(color = "grey", show.legend = FALSE) +
  scale_fill_manual(values = CPCOLS) +
  theme_minimal() +
  theme(
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text.y.left = element_text(size = 12, angle = 0, color = "gray72"),
    axis.title.y = element_text(color = "gray72")
  ) +
  xlab("number of goats tested on farm") + ylab("Farm Number") +
  ggtitle("Results of MAP testing")



# CWHP areas plot for limited farms
#plot_title_cwhp2 <- ggtitle("CWHP distribution (Selected Farms)")
# p_cwhp2 <- mcmc_areas(
#     posterior,
#     pars = vars(param_range("CWHP", rev(farm_ids2))),
#     area_method = "scaled height"
#   ) +
#   scale_y_discrete(expand = expansion(add = c(0, 0))) +
#   plot_title_cwhp2+
#   geom_vline(xintercept = .25) +
#   xlim(c(0, .3))

# 50% equal-tailed intervals + medians for HIP[i] (plot only; leaves your 90% table untouched)
df_cwhp50 <- fit5 %>%
  gather_draws(CWHP[i]) %>%
  median_qi(.width = 0.50)  # 50% interval

df_jwhp50 <- goat_sum %>%
  select(-i_s, -i_e) %>%      # same as you did earlier
  bind_cols(df_cwhp50) %>%
  rename(median_CWHP_50 = .value)

# Compact summary we’ll join to the density data

cwhp_summ50 <- df_jwhp50 %>%
  transmute(
    i,
    parameter = paste0("CWHP[", i, "]"),
    median    = median_CWHP_50,
    lower50   = .lower,
    upper50   = .upper
  )


cwhp_farm_lookup <- tibble(
  farm_number = farm_order,
  parameter   = paste0("CWHP[", farm_order, "]")
)

# density plot information
dens_df <- mcmc_areas_data(
  posterior,
  pars        = vars(param_range("CWHP", farm_order)),
  prob_outer  = 1,
  area_method = "equal area"
  ) %>%
  left_join(cwhp_farm_lookup, by = "parameter") %>%
  mutate(    farm_number = factor(farm_number, levels = farm_order)) %>%
  left_join(cwhp_summ50, by = "parameter")

# median value for each farm
facet_df <- dens_df %>%
  distinct(farm_number, median)


# labeller function to show median value in facet labels
median_labeller <- function(farm_values) {
  med <- facet_df$median[match(farm_values, facet_df$farm_number)]
  sprintf("med = %.2f", med)
}



my_cwhp_plot <- ggplot(dens_df, aes(x = x, y = density)) +
     geom_line(color = "gray30", linewidth = 0.6) +
     geom_ribbon(
          data = ~ dplyr::filter(.x, x >= lower50 & x <= upper50),
          aes(ymin = 0, ymax = density),
          fill = "#4C78A8",
          alpha = 0.35,
          color = NA
     ) +
     geom_vline(
          data = dplyr::distinct(dens_df, farm_number, median),
          aes(xintercept = median),
          color = "black",
          linewidth = 0.6,
          linetype = "dashed"
     ) +
     facet_wrap(
          vars(farm_number),
          ncol = 1,
          scales = "free_y",
          strip.position = "left",
          labeller = labeller(farm_number = median_labeller)
     ) +
     coord_cartesian(xlim = c(0, .32)) +
     labs(
          title = "CWHP distribution",
          subtitle = "Posterior with central 50% interval",
          x = "CWHP",
          y = NULL
     ) +
     theme_minimal(base_size = 12) +
     theme(
          strip.text.y.left = element_text(
               size = 10,
               angle = 0,
               hjust = 1,
               color = "gray40"
          ),
          strip.background = element_blank(),
          
          axis.text.y  = element_blank(),
          axis.ticks.y = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor = element_blank()
     )



p_res2 + my_cwhp_plot + plot_layout(widths = c(.8, 1.2))

# Combine plots
#p_cwhp2 + p_res2 + plot_layout(widths = c(1.5, 1))



#ggsave(filename=here("figs","cwhp_sel_goat2.png"), width=8,height=6,units="cm"   )
ggsave(filename=here("figs","cwhp_sel_goat2.tiff"), width=8,height=6,units="cm" ,dpi=300  )

```


#### Combined figure:

For the paper, I anticipate combining HIP and CWHP plots on this reduced dataset.


```{r}
# Define your farm list and get IDs
farm_list2 <- c(4, 11, 28, 5, 8, 2, 18, 13)
farm_ids2 <- goat_sum %>%
  filter(farm_number %in% farm_list2) %>%
  arrange(match(farm_number, farm_list2)) %>%
  pull(id)

# Build a named vector (if needed for axis labeling)
param_names2 <- rev(paste0("HIP[", farm_ids2, "]"))
farm_labels2 <- rev(paste0("Farm[", farm_list2, "]"))

# HIP plot for the selected farms
plot_title_hip2 <- ggtitle("HIP distribution (Selected Farms)")
p4g1_2 <- mcmc_areas(
    posterior,
    pars = vars(param_range("HIP", farm_ids2)),
    area_method = "scaled height"
  ) +
  plot_title_hip2

# CWHP plot for the selected farms
plot_title_cwhp2 <- ggtitle("CWHP distribution (Selected Farms)")
p4g3_2 <- mcmc_areas(
    posterior,
    pars = vars(param_range("CWHP", farm_ids2))
  ) +
  plot_title_cwhp2 +
  geom_vline(xintercept = .25) +
  xlim(c(0, .3))

# Tile plot for the selected farms
p4g2_2 <- g2 %>%
  filter(farm_number %in% farm_list2) %>%
  ggplot(aes(x = goat_num, y = fct_inorder(test), fill = result)) +
  facet_wrap(
    ~fct_relevel(as.factor(farm_number), as.character(farm_list2)),
    ncol = 1, strip.position = "left", dir = "v"
  ) +
  geom_tile(color = "grey", show.legend = FALSE) +
  scale_fill_manual(values = CPCOLS) +
  theme_minimal() +
  theme(
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text.y.left = element_text(size = 12, angle = 0, color = "gray72"),
    axis.title.y = element_text(color = "gray72")
  ) +
  xlab("number of goats tested on farm") + ylab("Farm Number") +
  ggtitle("Results of MAP testing (Selected Farms)")

# Combine the three plots
p4g1_2 + p4g2_2 + p4g3_2 + plot_layout(widths = c(1.5, 1.5, 1.5))
```


