---
title: "Using stan with heirarchical model"
subtitle: "Preliminary analysis of Sheep data"
author: "Joe Skufca"
date: '2022-01-19'
output:
  html_document:
    toc: yes
    df_print: paged
    code_folding: hide
  html_notebook:
    toc: yes
---

Experiment 8 provides an initial report on the "sheep" dataset, using an approach that equivalent to the approach of Baumen et al in the 2011 Ontario study, with some slight modifications:

* We use a two-latent class model, which assumes each animal is classes as either "infected" or "non-infected."
* We make estimates of within herd prevalence for each herd.


Additionally, this preliminary reporting uses "priors" parameterized as done in that 2011 paper, as modified slightly (as we await further updates:)

 


### Setup

We load anticipated packages and set rstan options.

```{r setup}
options(digits = 3)
knitr::opts_chunk$set( highlight=T,message=FALSE, warning=FALSE)

library(tidyverse)
library(patchwork)
library(janitor)
library(PriorGen)


library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores(logical = FALSE))

library(bayesplot)
library(here)
library(tidybayes)
library(shinystan)
library(DT)
```

## Preparing the sheep data

We will analyze the sheep data, but using a hierarchical model (with each farm treated as a herd).

### Data

Load data.  

I have made some choices about data removal (removing sheep with suspect elisa results, but we should discuss alternatives )

```{r message=FALSE, warning=FALSE, results="asis"}
sheep_file=here("data","sheep_individuals.xlsx")

sheep = readxl::read_excel(sheep_file, sheet = "Sheet1")%>%
     clean_names() %>%
     select(animal_number,farm_number,  elis_aresult, pc_rresult, map_result) %>%
     rename(elisa=elis_aresult,pcr=pc_rresult,culture=map_result) %>%
     filter(elisa  %in% c("p", "n")) %>%
     filter(culture %in% c("p", "n")) %>%
     mutate(across(elisa:culture, ~ (.x == "p") + 0)) 
gtsummary::tbl_summary(sheep %>% select(-animal_number))

```

### Assemble Stan model data

I include here a breakdown table by farm, as it is part of assembling the data structure required to input to the stan code. This table will help us to understand why certain farms are deemed to be higher likelihood of being infected as well as differing levels of within farm prevalence.

```{r}
sheep_sum=sheep %>% rowid_to_column("id") %>% group_by(farm_number) %>%
     summarise(i_s=first(id),i_e=last(id),n=n(),across(elisa:culture,sum))
sheep_sum 
```

For the hyperparamters we will need below, I work from data provided in Jessica's email dtd 20 Jan 2022 at 1611, where some modifications are required, in which case I will refer to Bauman 


For Elisa - use Bauman prior's - but should update based on their results.

For PCR Sens - use mode of 85, with 95% prob of above 20%

```{r}
findbeta(themode = .85,percentile.value = .2,lower.v = FALSE)
```

For Culture Sens - mode of 6%, 95% probability that it is less that 13%

```{r}
findbeta(themode = .06,percentile.value = .13,lower.v = TRUE)
```
For specificity:

ELISA: mode .99  but  95% sure that it's > .9
```{r}
findbeta(themode = .99,percentile.value = .9,lower.v = FALSE)
```

PCR: mode .978  but  95% sure that it's > .5   (some guessing here).
```{r}
findbeta(themode = .978 ,percentile.value = .5,lower.v = FALSE)
```

Culture: mode .995  but  95% sure that it's > .9   (some guessing here).
```{r}
findbeta(themode = .995 ,percentile.value = .9,lower.v = FALSE)
```


#### Help required with this prior

I filled in the numbers, but we would need expert opinion to pick quantities based on "prior" belief and literature review:


For HTP hyperparameters: We suspect that about **30%** of farms have Johnes present in the sheep flock, and are **95%** certain that the fraction of infected farms is less that **67%**.

```{r}
findbeta(themean = .30,percentile.value = .67,lower.v = TRUE)
```

To generate the hyperparameters for $\mu$ and $\psi$ so that we can prescribe priors for within herd prevalence, we complete the  portions of this sentence, where I have filled him with my guesses based on nothing but a swag to allow the computation to move forward:

> The mean prevalence of a disease/infection for the contaminated farms within an area/region
> is thought to be **0.10** and we are 95% confident that it is not more than **0.20.**
> Within this area, we are also confident that **90%** of all units have a prevalence
> less or equal to **0.35** and we are 95% certain that it does not exceed **0.45**.

```{r}
findbetamupsi(themean=0.10, percentile=0.95, percentile.value=0.20,
 lower.v=TRUE,
psi.percentile=0.90, percentile.median=0.35, percentile95value=0.45)
```




  
```{r}
# prior from Ontario 2011 paper
data5= list(F=21,
            I=301,
            I_s=sheep_sum$i_s,
            I_e=sheep_sum$i_e,
            J=3,
            y=as.matrix(sheep %>% select(3:5)),
            alpha_sens = c( 9.26,   2,    5.02),
            beta_sens =  c( 22.23,   1.18,    63.92),
            alpha_spec = c(34.17 , 4.5 , 31.08),
            beta_spec=   c( 1.34,   1.08,    1.15),
            htp_alpha=1.34,
            htp_beta=3.13,
            mu_alpha=3.09,
            mu_beta=27.29,
            psi_a=1.91,
            psi_b=.87
                    )

```


### fitting the model

This model takes a bit of time to run, depending upon the power of your machine.  On my office machine, with model compiled, I can compute a new simulation (new parameters) usually in about a minute.
```{r}
fit5=stan(file=here("src","heir_multi_test_B.stan"),data=data5,
          warmup=500,chains=12,iter=2000)
```

###  Visualizing results

Primarily, here, I will be showing results with graphs, using a bit of words to describe.  

The plots will all be posterior distribution of various parameter values.

```{r}
posterior=as.matrix(fit5)
```


### Herd True Prevalence

Herd True Prevalence (HTP) is the fraction of herd with at least one infected animal.

```{r}
plot_title <- ggtitle("Sheep Farm - Herd Level Prevalence",
                      "with medians and 80% intervals")
p1=mcmc_areas(posterior,prob=.8,pars="HTP")  + plot_title

p2=mcmc_dens_overlay(as.array(fit5),pars="HTP")+theme_minimal()
p1+p2
```

For a particular parameter, we can ask for detailed summary information in many different forms.  Below I provide a point estimate (median) along with an 90% hdi interval (highest density interval) estimate.   The median value indicates that about 1/3 of the farms are infected.
```{r}

fit5 %>% gather_draws(HTP) %>% median_hdi(.width=.9) %>% datatable %>% formatRound(2:4,digits=3)
```

As this mean value checks (somewhat) with the fraction of herds that had positive test results, we imagine that maybe the model is doing something that we expect.

### Sensitivity and Specificity

A critical goal to this grant, as well as to this analysis overall is ... how accurate are these tests.  Sensitivity and Specificity estimates are an *output* of the model, though we have discussed that they are affected by the assumed prior distributions.  For the analysis shown, the posterior distributions for these measures:

Specificity is very high, though ELISA is much worse than the others.

```{r }
mcmc_areas_ridges(posterior,regex_pars=c("spec"))+
     scale_y_discrete(labels=c("spec_elisa","spec_pcr","spec_cult"))
```

Sensitivity, however, is likely less than 0.5, and worst for pcr.

```{r}
mcmc_areas(posterior,regex_pars=c("sens"))+
     scale_y_discrete(labels=c("sens_elisa","sens_pcr","sens_cult"))
```
```{r}
fit5 %>% gather_draws(spec[i],sens[i]) %>% median_hdi(.width=.9) %>% 
     arrange(.variable) %>% datatable %>% formatRound(3:5,digits=3)
```

### Herd Infection Probability

The way I formulated the model allows me to compute a parameter which I call _herd infection probability_ (HIP), which give the probability that a particular herd has at least one infected animal in the group of animals tested.  (Note the care of wording.  The model is not aware of the total number of animals on the farm, which would be necessary to actually perform this calculation.)

```{r fig.height=10}
mcmc_areas(posterior,pars=vars(param_range("HIP",c(1:21))))
```
```{r}
fit5 %>% gather_draws(HIP[i]) %>% median_hdci(.width=.9) %>% datatable() %>% formatRound(3:5,digits=3)
```

It is informative to merge these median values with the farm-level test results to the implications.

```{r}
df_h=fit5 %>% gather_draws(HIP[i]) %>% median_hdci(.width=.9) %>% select(i,.value)
df_j=sheep_sum %>% select(-i_s,-i_e) %>% rowid_to_column() %>%
     left_join(df_h,by=c("rowid"="i")) %>% rename(median_HID=.value)

df_j %>% datatable %>% formatRound(7,digits=3)
```

### Farm Level Prevalance

We are also interested in understanding, _under the condition that a herd is infected, what is the prevalence of infection within that herd_ (CWHP).  Stated another way, what fraction of a herd is infected (when that number is not 0).

Recognizing that we have pretty sparse data for any particular herd (small sample size for the herd, small number of positives), the wide variability in the parameter estimate is not surprising.   However, we do see strong consistency if we focus on (median levels) especially if we look at herds that are likely infected.

If CWHP is relatively low, it means the disease is does not spread completely throught the herd.  *This phenomena needs to be better understood.*

```{r}
mcmc_areas(posterior,regex_pars=c("CWHP")) 
```

```{r}
fit5 %>% gather_draws(CWHP[i]) %>% median_hdci(.width=.9) %>% datatable %>% formatRound(3:5,digits=3)
```



## Is our model any good?

One way to address this question is to see if our model (with the fitted parameters) produces data "like" our original.  I will discuss this in more detail at our meeting.


```{r}
df_a=sheep %>% count(elisa,pcr,culture) %>% arrange(elisa,pcr,culture)
df_a
```

Let's create a helper function that takes a sampled parameter draw and generates a replicate sample:

```{r}
this_samp=2222
make_rep <- function(this_samp) {
  A=posterior[this_samp,]
  
  pi0=rbernoulli(21,A["HTP"])
  
  cwhp=A[1:21]
  spec=A[c("spec[1]","spec[2]","spec[3]")]
  sens=A[c("sens[1]","sens[2]","sens[3]")]
  
  #pi1=(pi0) %*% t(1-spec) 
  
  
  
  y_rep=matrix(,nrow=301,ncol=3)
    for (n in 1:21) {
  
      for (k in data5$I_s[n]:data5$I_e[n]) {
           # is this sheep infected determines value for pi1
           if (!pi0[n]) {  # herd not infected
                pi1=1-spec
           } else if (rbernoulli(1,cwhp[n])) { #sheep infected
                pi1=sens
           } else {
               pi1=1-spec 
           }
           
           
           
           y_rep[k,1:3]=rbernoulli(3,pi1)
           }
      }
  
  y_rep=as.tibble(y_rep+0) %>% rename(elisa=V1,pcr=V2,culture=V3)
}
```

```{r}

make_rep(sample(5000,1)) %>%  count(elisa,pcr,culture) %>% arrange(elisa,pcr,culture)
```





