---
title: "Using stan with heirarchical model"
author: "Joe Skufca"
date:  "2021-12-23"
output: 
    html_notebook:
       toc: true
---

This experiment intends to analyze data from each farm as a herd, with common specificity and sensititive.  We assume on a single test (for which I will use the culture data)    The dataset of focus here is the "goat" dataset.

### Setup

```{r setup}
options(digits = 3)
knitr::opts_chunk$set( fig.width = 6,highlight=T)

library(tidyverse)
library(janitor)
library(PriorGen)


library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores(logical = FALSE))

library(bayesplot)
library(here)
library(tidybayes)
```

## Example 1

We will analyze the goat data, but without a heirarchical model (so no need to include which farm).

### Data

Load data.  We eliminate the one sample that was questionable.


```{r}
goat_file=here("data","goat_results.xlsx")

goat = readxl::read_excel(goat_file, sheet = "Sheet1") %>%
     clean_names() %>%
     select(sample_number,farm_number,  elisa, pcr, culture) %>%
     filter(elisa  %in% c("p", "n")) %>%
     filter(culture %in% c("p", "n")) %>%
     mutate(across(elisa:culture, ~ (.x == "p") + 0)) 


```

### Assemble Stan model data

In addition to the data, we need information for prior assumptions regarding sensitivity and specificity.

```{r}
goat_sum=goat %>% rowid_to_column("id") %>% group_by(farm_number) %>%
     summarise(i_s=first(id),i_e=last(id))
```




```{r}
data1= list(F=27,
            I=316,
            I_s=goat_sum$i_s,
            I_e=goat_sum$i_e,
            y=goat$culture,
            alpha_sens = 5,
            beta_sens = 7,
            alpha_spec = 1500,
            beta_spec= 7
                    )

# int<lower=1> F; //number of farms
#   int<lower=1> I; //total number of animals tested
#   int<lower=1> I_s[F]; //starting index for animals from from f
#   int<lower=1> I_e[F]; //ending index for animals from from f
# 
#   int<lower=1,upper=2> y[I]; // test resuls  0 for neg, 1 for pos
#   // params for prior for sensitivities and specificities
#   real<lower=0> alpha_sens;
#   real<lower=0> beta_sens;
#   real<lower=0> alpha_spec;
#   real<lower=0> beta_spec;
```


### fitting the model

```{r}
fit1=stan(file=here("src","heir_single_test.stan"),data=data1,
          warmup=500,chains=12,iter=3000)
```

###  Visualizing results

```{r}
posterior=as.matrix(fit1)
```


Prevalence

```{r}
plot_title <- ggtitle("Posterior distributions",
                      "with medians and 80% intervals")
mcmc_areas(posterior,prob=.8,pars="HTP")


```

```{r}
summary(fit1)
```

### Which farms?

```{r}
mcmc_areas_ridges(posterior,regex_pars=c("CWHP"))
```
### Farm Level prevelence

```{r}
mcmc_areas_ridges(posterior,regex_pars=c("HID\\[1[0-9],1"))
```


### Chain mixing

```{r}
plot(fit1, plotfun = "trace", pars = c("beta1"), inc_warmup = TRUE)
plot(fit1, plotfun = "rhat")
```


```{r}

mcmc_intervals(posterior,regex_pars=c("eta"))

```



########################








