---
title: "Using stan without heirarchical model"
author: "Joe Skufca"
date:  "2021-12-23"
output: 
    html_notebook:
       toc: true
---

This experiment intends to use a commands from the `rater` package to analyze data for classes.  It likely provides less direct information for sensitivity and specificity, but will predict class for each animal.

### Setup

```{r setup}
options(digits = 3)
knitr::opts_chunk$set( fig.width = 6,highlight=T)

library(tidyverse)
library(janitor)


library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores(logical = FALSE))

library(bayesplot)
library(here)
```

## Example 1

We will analyze the goat data, but without a heirarchical model (so no need to include which farm).

### Data

Load data


```{r}
goat_file=here("data","goat_results.xlsx")

goat = readxl::read_excel(goat_file, sheet = "Sheet1") %>%
     clean_names() %>% 
     select(sample_number,elisa,pcr,culture) %>%
     filter(elisa  %in% c("p","n")) %>%
     filter(culture %in% c("p","n")) %>% 
     mutate(across(elisa:culture, ~ (.x == "p")+1)) 


```

### Assemble Stan model data

For multi_test_spec_sens_prev.stan:

>   int<lower=1> I; //number of animals tested
>   int<lower=1> J; // number of tests - 3 for our work
>   
>      
>   int<lower=0,upper=1> y[I,J]; // test resuls  1/0
> 
>   // params for prior
>   row_vector<lower=0>[J] alpha_sens;
>   row_vector<lower=0>[J] beta_sens;
>   row_vector<lower=0>[J] alpha_spec;
>   row_vector<lower=0>[J] beta_spec;



```{r}
data2= list(I=316,
            J=3,
            y=as.matrix(goat %>% select(2:4)),
            alpha_sens = c(167,167,167),
            beta_sens = c(43,43,43),
            alpha_spec = c(43,43,43),
            beta_spec= c(5.62,5.62,5.62)
                    )
```


### fitting the model

```{r}
fit2=stan(file=here("src","multi_test_spec_sens_prev.stan"),data=data2)
```

###  results

```{r}
df_sum=summary(fit2)
```





