---
title: "Simple prevelance/sensitivity/specificity experiment"
author: "Joe Skufca"
date:  "2021-12-17"
output: 
    html_notebook:
       toc: true
---

This experiment intends to use a simple stan model that will give prevelance, specificiy, and sensitivity bayesian parameter analysis on a model of a single site collection.  Prior (for sens and spec) will come from a beta distribution.

### Setup

```{r setup}

options(digits = 3)
knitr::opts_chunk$set( fig.width = 6,highlight=T)

library(tidyverse)
library(janitor)

library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores(logical = FALSE))

library(bayesplot)
library(here)
```

## Example 1

All will use the same compiled model but we will perhaps look at a few different data sets.

### Data

Coefficients for priors are "thin-air" but use `PriorGen` package to have interpretable parameters.

To cite package ‘PriorGen’ in publications use:

  Polychronis Kostoulas (2018). PriorGen: Generates Prior Distributions
  for Proportions. R package version 1.1.2.
  https://CRAN.R-project.org/package=PriorGen

```{r}
PriorGen::findbeta(themode = .8, percentile.value = .75)
```


```{r}
PriorGen::findbeta(themode = .9, percentile.value = .8)
```


  int y_sample; // number of positives
  int n_sample; // number of samples
  // params for prior
  real<lower=0> alpha_sens;
  real<lower=0> beta_sens;
  real<lower=0> alpha_spec;
  real<lower=0> beta_spec;


```{r}
data1= list(y_sample = 20,
            n_sample=297,
            alpha_sens = 167,
            beta_sens = 43,
            alpha_spec = 43,
            beta_spec= 5.62
                    )
```

### Fit the model

```{r}
fit1=stan(file=here("src","simple_spec_sens_prev.stan"),data=data1)
```

### visualize results

```{r}
posterior <- as.array(fit1)
mcmc_dens(
  posterior
)
```

## Conclusion

Without "ground truth" or "gold-standard" with regard to the true state of infection, the analysis cannot provide concensus.  The bi-modal solution for prevelance is a key to understanding.

Either:

* low prevalence and reasonable accurate tests, or
* high prevalence, and tests with poor performance (especially with regards to sensitivity.)