---
title: "Experiment 1: mapping"
author: "Joe Skufca"
date:  "2021-10-26"
output: html_notebook
---

A first attempt to map the farm locations.


```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 6
)
``` 
## Packages

```{r}

library(tmap)
library(tmaptools)
library(sf)
library(leaflet)
library(readxl)
library(janitor)

library(tidyverse)
```



## Load data

From the file provided by Cornell:

```{r}
file_farms=here::here("data","SHEEP demographics with  NO NAMES.xlsx")
df1=read_excel(file_farms) %>% clean_names()  %>% drop_na(address) %>%
  mutate(sampled=case_when(
    sampled=="YES" ~ 1,
    sampled=="N" ~ 0
  ), johnes=str_to_lower(johnes))


```




### Geocoding with `hereR`

```{r}
library(hereR)
set_key("3AbZA9I7RALc6NGEqy5jX9HXyW0_pciFFwrS8DF1s6M")

```



```{r}
dfh=geocode(df1$address)
```

Since df1 and dfh have sam number of rows, we may join into a single `sf` object.

```{r}
df_j=bind_cols(dfh,df1)

```

## Let's use tmap to build a bubble map on the joined dataset, with a leaflet background.


```{r, results='asis'}
tmap_mode("view")
tm_shape(df_j)+tm_symbols("total_herd_size",
                          col="johnes",
                          id="total_herd_size")
```


## Let's visualize by town

```{r}
library(tidycensus)
```

Get county data for NYS and spatially join the farm data:

```{r include=FALSE}
#ny <-tigris::counties(state="NY")
ny=tigris::county_subdivisions('NY')
ny=st_transform(ny,4326)
ny_j=st_join(ny,df_j,left=TRUE) %>% drop_na(total_herd_size)

```

Let's try to use dplyr methods to aggregate and visualize total count of farms:

```{r, results='asis'}
df4=ny_j %>% group_by(NAME) %>% summarise(farms=n(),
                                          total_herd=sum(total_herd_size),
                                          mean_exper=mean(years_expierence)) 

tm_shape(df4) +tm_polygons(col=c("total_herd","farms","mean_exper"),id="NAME") +tm_facets(ncol=3)
```





